### 2023-03-02 Epidemics_submission/outbreak_response_models_review_v2.tex

A little over half of the human disease papers used observed data for parametrization or validation. Additionally, only 4 studies shared their code in an easy to access form. On the other hand, the FMD models were generally not validated with observed data. Owing to this, it might be difficult to reproduce some of the results in these outbreak response modelling studies. We recognise that data sharing in public health raises a lot of debate regarding privacy and intellectual property, and often, authors are hindered by institutional data sharing policies \cite{Kim2016}. We, however, recommend that data and code be shared where possible to promote open science practices that help advance the field.

This review had several limitations. First, outbreak response models and analyses are not always published in peer-reviewed journals, but this systematic review only focussed on peer-reviewed articles. It is, therefore, possible that some relevant studies were missed by our search strategy. Second, we used the author list and affiliations to classify studies as either purely academic or mixed. However, this could cause some mixed collaborations to be misclassified as purely academic, especially in cases where non-academics contributed to papers classified as purely academic but were not included on the author list. It is, however, standard practice to include individuals who contributed substantially to a piece of scientific writing, hence, if that was not done for a paper, it could imply that the level of interaction was not high enough to warrant the credit of authorship. Third, in the absence of an explicit measure of contribution of mathematical modelling to outbreak response decision-making, we used mixed collaborations as a proxy. Thus, we may be under-estimating the number of mixed collaborations in the literature. Forth, we only surveyed the literature on a specific study design – mechanistic models. We excluded statistical models and other computational models, which are not mechanistic. Hence, the results of this systematic review should only be interpreted in the context of the mechanistic modelling landscape. Furthermore, it is also possible that some purely academic collaborations contribute to decision-making whereas some mixed collaborations are purely an academic exercise. Also, we  conducted our database searches in January 2020 and therefore do not reflect the literature, or any changes in practice, associated with the COVID-19 pandemic. Lastly, by only including papers published in English, we most certainly missed papers published in other languages.

Numerous factors could explain the patterns we have observed in this review, and we would recommend future studies that will use appropriate study designs, for example, interviews of modelling groups and public health decision-makers, to explain why certain model choices and modelling practices were made by the collaboration types. Future studies should investigate when modelling results directly impacted decision-making and what determined that to identify best practices that will strengthen the link between modelling and decision-making in the future. We expect that collaboration between academia, decision-makers, and local experts will enhance decision-making by accounting for aspects of policy and decision-making that might be overlooked in studies conducted mainly as a theoretical exercise.

\subsection*{Other information}
A protocol following the PRISMA guidelines for systematic review protocols and outlining the procedures for this systematic review was registered on PROSPERO with registration number CRD42020160803 and published through a peer-reviewed process~\cite{Azam2020}.

\section*{Supporting information}

\paragraph*{S1 File}
\label{S1_File}
{\bf Search strings and results, data items extracted, and questionnaire for data extraction.}

\paragraph*{S1 Fig}
\label{S1_Fig}
{\bf Collaboration patterns in time (proportions per year).}

\paragraph*{S2 File}
\label{S2_File}
{\bf Analysis of extracted data.}

\paragraph*{S3 Table}
\label{S3_Table}
{\bf PRISMA checklist.}

%\section*{Acknowledgments}

% Use the PLoS provided BiBTeX style
\bibliography{references.bib}
\bibliographystyle{vancouver}
\nolinenumbers
\end{document}
